---
title: "HW1_analysis"
author: "AnnaClaire Marley"
date: "4/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(dataRetrieval)
library(Kendall)
library(gridExtra)
```

```{r, echo = FALSE}
# source all my functions
source('../HW_1/functions/read_chunks.R')
source('../HW_1/functions/extract_months.R')
source('../HW_1/functions/graph_season_trend.R')
source('../HW_1/functions/plot_month_anomaly.R')
```

# Precipitation Analysis
```{r, include=FALSE}
precip <- read_chunks("data")
```

#### Raw precipitation data from 1981 to February 2020
```{r, echo = FALSE}
ggplot(data = precip, aes(x = date, y = precipitation)) +
  geom_line()
```

#### Average yearly precipitation from Jan 1 1981 to Dec 31 2019

```{r, echo = FALSE}
anuual_av <- precip %>% 
  arrange(date) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(precip_av = mean(precipitation)) %>% 
  filter(year != 2020)

ggplot(anuual_av, aes(x = year, y = precip_av)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Annual Average Precipitation"
  ) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()
```

```{r, echo = FALSE}
# calculate linear model

precip_av_lm <- lm(precip_av~year, anuual_av)
precip_av_lm_sum <- summary(precip_av_lm)

```

Precipitation is decreasing by -0.003404 mm/yr and is statistically signficant, pvalue = 0.02



```{r, echo = FALSE}

#### Average spring precipitation from Jan 1 1981 to Dec 31 2019
# only look at spring precipitation 
spring_precip <- extract_months(precip, wanted_months = c(3,4,5))

# create dataframe of average yearly spring precipitation
spring_av <- spring_precip %>% 
  arrange(date) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(precip_av = mean(precipitation)) %>% 
  filter(year != 2020)

# graph it and add trendline 
spring_precip_plot <- graph_season_trend(spring_av, 
                   climate_data = spring_av$precip_av,
                   title = "Spring",
                   ylabel = "Average Precipitation (mm)")

```


```{r, echo = FALSE}
#### Average summer precipitation
# only look at summer precipitation 
summer_precip <- extract_months(precip, wanted_months = c(6,7,8))

# create dataframe of average yearly summer precipitation
summer_av <- summer_precip %>% 
  arrange(date) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(precip_av = mean(precipitation)) %>% 
  filter(year != 2020)

# graph it and add trendline 
summer_precip_plot <- graph_season_trend(summer_av, 
                   climate_data = summer_av$precip_av,
                   title = "Summer",
                   ylabel = "Average Precipitation (mm)")

```



```{r, echo = FALSE}
#### Average Fall precipitation 
# only look at fall precipitation 
fall_precip <- extract_months(precip, wanted_months = c(9,10,11))

# create dataframe of average yearly fall precipitation
fall_av <- fall_precip %>% 
  arrange(date) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(precip_av = mean(precipitation)) %>% 
  filter(year != 2020)

# graph it and add trendline 
fall_precip_plot <- graph_season_trend(fall_av, 
                   climate_data = fall_av$precip_av,
                   title = "Fall",
                   ylabel = "Average Precipitation (mm)")

```




```{r, echo = FALSE}
#### Average winter precipitation by water year
winter_precip <- extract_months(precip, wanted_months = c(12,1,2))

winter_av <- winter_precip %>% 
  rename(Date = date) %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  summarize(precip_av = mean(precipitation)) 

# graph it and add trendline 
wint_precip_plot <- ggplot(winter_av, aes(x = waterYear, y = precip_av)) +
                  geom_line() +
                  labs(
                    title = "Winter",
                    x = "Year",
                    y = "Average Precipitation (mm)"
                  ) +
                  geom_smooth(method = "lm", se = FALSE) +
                  scale_x_continuous(expand = c(0,0)) +
                  theme_bw()

```

```{r, echo = FALSE}
# grid all the season graphs into one graph
grid.arrange(spring_precip_plot, summer_precip_plot, 
             fall_precip_plot, wint_precip_plot, ncol = 2)
```


#### Evaluate strengths of the trends
```{r, echo = FALSE}
# annual average precipitation
av_precip_mk <- MannKendall(anuual_av$precip_av)
# average spring precip
spring_precip_mk <- MannKendall(spring_av$precip_av)
# average summer precip
sum_precip_mk <- MannKendall(summer_av$precip_av)
# average fall precip
fall_precip_mk <- MannKendall(fall_av$precip_av)
# average winter precip in water years
win_precip_mk <- MannKendall(winter_av$precip_av)

```
- Average annual precipitation: tau = `r round(av_precip_mk$tau, 2)`, pvalue = `r round(av_precip_mk$sl, 2)`
- Average spring precipitation: tau = `r round(spring_precip_mk$tau, 2)`, pvalue = `r round(spring_precip_mk$sl, 2)`
- Average summer precipitation: tau = `r round(sum_precip_mk$tau, 2)`, pvalue = `r round(sum_precip_mk$sl, 2)`
- Average fall precipitation: tau = `r round(fall_precip_mk$tau, 2)`, pvalue = `r round(fall_precip_mk$sl, 2)`
- Average winter precipitation: tau = `r round(win_precip_mk$tau, 2)`, pvalue = `r round(win_precip_mk$sl, 2)`

#### Winter precipitation including March
```{r, echo = FALSE}
#### Average winter precipitation by water year
winter_precip_m <- extract_months(precip, wanted_months = c(12,1,2,3))

winter_av_m <- winter_precip_m %>% 
  rename(Date = date) %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  summarize(precip_av = mean(precipitation)) 

# graph it and add trendline 
ggplot(winter_av_m, aes(x = waterYear, y = precip_av)) +
      geom_line() +
      labs(
        x = "Year",
        y = "Average Precipitation (mm)"
      ) +
      geom_smooth(method = "lm", se = FALSE) +
     scale_x_continuous(breaks = seq(1981,2021, by = 3), expand = c(0,0)) +
      theme_bw(base_size = 15)

```

```{r, echo = FALSE}
### stats on winter precip with march ###

# mann kendall
win_precip_mk_m <- MannKendall(winter_av_m$precip_av)

# linear model

summary(lm(precip_av ~ waterYear, data = winter_av_m))
```



#### Anomaly of winter months
```{r, echo = FALSE}
dec_anom_plot <- plot_month_anomaly(precip)
jan_anom_plot <- plot_month_anomaly(precip, months = c(1), anom_year = 1991, title = "January")
feb_anom_plot <- plot_month_anomaly(precip, months = c(2), anom_year = 1991, title = "February")
mar_anom_plot <- plot_month_anomaly(precip, months = c(3), anom_year = 1991, title = "March")
grid.arrange(dec_anom_plot, jan_anom_plot, feb_anom_plot, mar_anom_plot, ncol = 2)
```


# Temperature Analysis

```{r, include=FALSE}
# read in datasets
tmax <- read_chunks("data", "*max_temp_*", variable = "temperature")
tmin <- read_chunks("data", "*min_temp_*", variable = "temperature")
```


### Raw maximum temperature data from 1980-2020
```{r, echo = FALSE}
ggplot(data = tmax, aes(x = date, y = temperature)) +
  geom_line()+
  labs(x = "Date",
       y = "Max Temperature (degreees C)")
```

### Raw minimum temperature data from 1980-2020

```{r, echo = FALSE}
ggplot(data = tmin, aes(x = date, y = temperature)) +
  geom_line()+
  labs(x = "Date",
       y = "Min Temperature (degreees C)")
```

#### Average annual maximum temperature
```{r, echo = FALSE}
anuual_av_temp_max <- tmax %>% 
  arrange(date) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(temp_av = mean(temperature)) 

graph_season_trend(anuual_av_temp_max, 
                   climate_data = anuual_av_temp_max$temp_av,
                   ylabel = "Average Annual Max Temperature (degrees C)")
```

#### Average annual minimum temperature
```{r, echo = FALSE}
anuual_av_temp <- tmin %>% 
  arrange(date) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(temp_av = mean(temperature)) 

graph_season_trend(anuual_av_temp, 
                   climate_data = anuual_av_temp$temp_av,
                   ylabel = "Average Annual Min Temperature (degrees C)")
```

#### Winter minimum temperatures 
```{r, echo = FALSE}
#### Average winter temperature by water year
winter_temp_min <- extract_months(tmin, wanted_months = c(12,1,2,3))

winter_av_temp_min <- winter_temp_min %>% 
  rename(Date = date) %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  summarize(temp_av = mean(temperature)) 

# graph it and add trendline 
ggplot(winter_av_temp_min, aes(x = waterYear, y = temp_av)) +
    geom_line() +
    labs(
      x = "Year",
      y = "Average Minimum Temperature (°C)"
    ) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(breaks = seq(1981,2021, by = 3), expand = c(0,0)) +
    theme_bw(base_size = 15)

```

```{r, echo = FALSE}
# stats for winter min tempeeartures

summary(lm(temp_av ~ waterYear, data = winter_av_temp_min))

# mann kendall
 MannKendall(winter_av_temp_min$temp_av)

```



#### Number of freezing days (temp below 0 degrees C)

```{r, echo = FALSE}
num_freeze <- tmin %>% 
  mutate(freeze = ifelse(temperature <= 0, 1, 0)) %>% 
  group_by(year = year(date)) %>% 
  summarize(ndayfr=sum(freeze)) %>% 
  filter(year != 2020)

graph_season_trend(num_freeze, climate_data = num_freeze$ndayfr, 
                   ylabel = "Number of Freezing Days")
```

Freezing days in the winter by water year
```{r, echo = FALSE}
# by water year
num_freeze_wy <- tmin %>% 
  extract_months(., wanted_months = c(12,1,2,3)) %>% 
  mutate(freeze = ifelse(temperature <= 0, 1, 0)) %>% 
  rename(Date = date) %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>%
  summarize(ndayfr=sum(freeze)) %>% 
  filter(waterYear != 1981)

# graph it and add trendline 
ggplot(num_freeze_wy, aes(x = waterYear, y = ndayfr)) +
    geom_line() +
    labs(
      x = "Year",
      y = "Number of Freezing days "
    ) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(breaks = seq(1980,2020, by = 3), expand = c(0,0)) +
    theme_bw(base_size = 15)
```

```{r, echo = FALSE}
# stats on freezing days
summary(lm(ndayfr ~ waterYear, data = num_freeze_wy))

```



#### Hottest day of the year temperature

```{r, echo = FALSE}

hotday <- tmax %>% 
  group_by(year = year(date)) %>% 
  summarize(hot_temp = max(temperature)) %>% 
  filter(year != 2020)

graph_season_trend(hotday, climate_data = hotday$hot_temp, 
                   ylabel = "Temperature (Degrees C)")

```



#### Winter max temperatures 
```{r, echo = FALSE}
#### Average winter temperature by water year
winter_temp_max <- extract_months(tmax, wanted_months = c(12,1,2))

winter_av_temp_max <- winter_temp_max %>% 
  rename(Date = date) %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  summarize(temp_av = mean(temperature)) 

# graph it and add trendline 
ggplot(winter_av_temp_max, aes(x = waterYear, y = temp_av)) +
    geom_line() +
    labs(
      title = "Winter",
      x = "Year",
      y = "Average Maximum Temperature (degrees C)"
    ) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(expand = c(0,0)) +
    theme_bw()

```

#### Hottest day of winter temperature by water year
```{r,echo = FALSE}
hotday_wint <- tmax %>% 
  extract_months(., wanted_months = c(12,1,2,3)) %>% 
  rename(Date = date) %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>%  
  summarize(hot_temp = max(temperature))

# graph it
ggplot(hotday_wint, aes(x = waterYear, y = hot_temp)) +
    geom_line() +
    labs(
      x = "Year",
      y = "Temperature (°C)"
    ) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(breaks = seq(1980,2020, by = 3), expand = c(0,0)) +
    theme_bw(base_size = 15)
```

hottest day of witner stats
```{r, echo = FALSE}
# hottest day of witner stats
summary(lm(hot_temp ~ waterYear, data = hotday_wint))

```

