---
title: "Projected Avocado Yields"
author: "AnnaClaire Marley"
date: "5/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(lubridate)
library(tidyverse)
library(stats)
library(tidyquant)


```


This markdown explores avocado yields in Ventura county under different climate projections.

Avocado yield equation (from Lobell et al., 2006):

Y = 17.71T<sub>x,-8</sub> - 0.29T<sup>2</sup><sub>x,-8</sub> + 3.25T<sub>n,5</sub> - 0.14T<sup>2</sup><sub>n,5</sub> + 1.00P<sub>-10</sub> - 0.31P<sup>2</sup><sub>-10</sub> - 288.09

- Y: yield anomaly (ton/acre)
- Tn: min temperature (degrees C)
- Tx: maximum temperature
- P: precipitation (mm)
- subscripted numbers represent month of climate variable with negative numbers repreesenting month from the year prior to harvest


## ACCESS 01

### RCP 4.5

Read in and clean data for climate projections
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# source data cleaning function
source('~/Documents/github/climate_adapt/HW_3/functions/clean_clim.R')

### max temperature ###
tmax_4.5_og <- read_csv("../HW_3/data/tasmax_day_ACCESS1-0_rcp45.csv") 

# clean and summarize by correct month
tmax_4.5 <- clean_clim(tmax_4.5_og, type = "tmax")

### min temp ###
tmin_4.5_og <- read_csv("../HW_3/data/tasmin_day_ACCESS1-0_rcp45.csv")

# clean and summarize by correct month
tmin_4.5 <- clean_clim(tmin_4.5_og, type = "tmin")

### precip ###
precip_4.5_og <- read_csv("../HW_3/data/pr_day_ACCESS1-0_rcp45.csv",
                          col_names=c("time", "precip"), skip = 1)

# clean and summarize by correct month
precip_4.5 <- clean_clim(precip_4.5_og, type = "precip")

```

Avocado anomaly 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# combine into one dataframe

RCP_4.5 <- left_join(tmax_4.5, tmin_4.5, by = "year") %>% 
  left_join(precip_4.5, by = "year")

# plot temp
ggplot(RCP_4.5, aes(x = year, y = tmax))+
  geom_line() +
  geom_line(aes(y = tmin))
# plot precip
ggplot(RCP_4.5, aes(x = year, y = precip))+
  geom_line() 

# avocado yield anomaly

yield_4.5 <- RCP_4.5 %>% 
  mutate(yield_anomaly = (17.71*lag(tmax,1) - 
                          0.29*lag(tmax,1)^2 + 
                          3.25*tmin - 
                          0.14*tmin^2 + 
                          1.00*lag(precip, 1) - 
                          0.31*lag(precip,1)^2 - 
                          288.09))
# plot it
ggplot(yield_4.5, aes(x = year, y = yield_anomaly)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Yield Anomaly (ton/acre)"
  ) +
  theme_classic()
```


### RCP 8.5

```{r, echo=FALSE, message=FALSE, warning=FALSE}

### tmax ###
tmax_8.5_og <- read_csv("../HW_3/data/tasmax_day_ACCESS1-0_rcp85.csv")

# clean
tmax_8.5 <- clean_clim(tmax_8.5_og, type = "tmax")

### tmin ###
tmin_8.5_og <- read_csv("../HW_3/data/tasmin_day_ACCESS1-0_rcp85.csv")

# clean
tmin_8.5 <- clean_clim(tmin_8.5_og, type = "tmin")

### precip ###
precip_8.5_og <- read_csv("../HW_3/data/pr_day_ACCESS1-0_rcp85.csv", 
                         col_names=c("time", "precip"), skip = 1)

# clean 
precip_8.5 <- clean_clim(precip_8.5_og, type = "precip")

```

Avocado anomaly 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# combine into one dataframe

RCP_8.5 <- left_join(tmax_8.5, tmin_8.5, by = "year") %>% 
  left_join(precip_8.5, by = "year")

# plot temp
ggplot(RCP_8.5, aes(x = year, y = tmax))+
  geom_line() +
  geom_line(aes(y = tmin))
# plot precip
ggplot(RCP_8.5, aes(x = year, y = precip))+
  geom_line() 

# avocado yield anomaly

yield_8.5 <- RCP_8.5 %>% 
  mutate(yield_anomaly = (17.71*lag(tmax,1) - 
                          0.29*lag(tmax,1)^2 + 
                          3.25*tmin - 
                          0.14*tmin^2 + 
                          1.00*lag(precip, 1) - 
                          0.31*lag(precip,1)^2 - 
                          288.09))
# plot it
ggplot(yield_8.5, aes(x = year, y = yield_anomaly)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Yield Anomaly (ton/acre)"
  ) +
  theme_classic()
```

ACCESS

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# graph together
ACCESS_yields <- yield_8.5 %>% 
  rename("ACCESS1-0 RCP 8.5" = yield_anomaly) %>% 
  left_join(yield_4.5, by = "year") %>% 
  dplyr::select(year, yield_anomaly, "ACCESS1-0 RCP 8.5") %>% 
  rename("ACCESS1-0 RCP 4.5" = yield_anomaly) 

ACCESS_gather <- ACCESS_yields %>% 
  gather(key = "projection", value = "anomaly", -year)

# plot
ggplot(ACCESS_gather, aes(x = year, y = anomaly)) +
  geom_line(aes(color = projection)) +
  geom_smooth(aes(color = projection), se = FALSE) +
  labs(
    x = "Year",
    y = "Yield Anomaly (tons/acre)",
    color = "Climate Projections"
  ) + 
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() 


```


## CCSM4

### RCP 4.5

```{r, echo=FALSE, message=FALSE, warning=FALSE}

### max temperature ###
tmax_4.5_og <- read_csv("../HW_3/data/tasmax_day_CCSM4_rcp45.csv")

# clean and summarize by correct month
tmax_4.5 <- tmax_4.5_og %>% 
  rename(tmax = "tasmax_day_CCSM4_rcp45") %>% 
  mutate(year = year(time)) %>% 
  mutate(month = month(time)) %>% 
  group_by(year, month) %>% 
  summarize(tmax = mean(tmax)) %>% 
  filter(month == 8) %>% 
  dplyr::select(year, tmax) %>% 
  mutate(tmax = tmax - 273.15) %>% 
  ungroup()

### min temp ###
tmin_4.5_og <- read_csv("../HW_3/data/tasmin_day_CCSM4_rcp45.csv")

# clean and summarize by correct month
tmin_4.5 <- tmin_4.5_og %>% 
  rename(tmin = "tasmin_day_CCSM4_rcp45") %>% 
  mutate(year = year(time)) %>% 
  mutate(month = month(time)) %>% 
  group_by(year, month) %>% 
  summarize(tmin = mean(tmin)) %>% 
  filter(month == 5) %>% 
  dplyr::select(year, tmin) %>% 
  mutate(tmin = tmin - 273.15) %>% 
  ungroup()

### precip ###
precip_4.5_og <- read_csv("../HW_3/data/pr_day_CCSM4_rcp45.csv",
                          col_names=c("time", "precip"), skip = 1)

# clean and summarize by correct month
precip_4.5 <- clean_clim(precip_4.5_og, type = "precip")

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# combine into one dataframe

CCSM4_RCP_4.5 <- left_join(tmax_4.5, tmin_4.5, by = "year") %>% 
  left_join(precip_4.5, by = "year")

# plot temp
ggplot(CCSM4_RCP_4.5, aes(x = year, y = tmax))+
  geom_line() +
  geom_line(aes(y = tmin))
# plot precip
ggplot(CCSM4_RCP_4.5, aes(x = year, y = precip))+
  geom_line() 


# avocado yield anomaly

CCSM4_yield_4.5 <- CCSM4_RCP_4.5 %>% 
  mutate(yield_anomaly = (17.71*lag(tmax,1) - 
                          0.29*lag(tmax,1)^2 + 
                          3.25*tmin - 
                          0.14*tmin^2 + 
                          1.00*stats::lag(precip,1) - 
                          0.31*stats::lag(precip,1)^2 - 
                          288.09))
# plot it
ggplot(CCSM4_yield_4.5, aes(x = year, y = yield_anomaly)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Yield Anomaly (ton/acre)"
  ) +
  theme_classic()
```


### RCP 8.5

```{r, echo=FALSE, message=FALSE, warning=FALSE}

### max temperature ###
tmax_8.5_og <- read_csv("../HW_3/data/tasmax_day_CCSM4_rcp85.csv")

# clean and summarize by correct month
tmax_8.5 <- tmax_8.5_og %>% 
  rename(tmax = "tasmax_day_CCSM4_rcp85") %>% 
  mutate(year = year(time)) %>% 
  mutate(month = month(time)) %>% 
  group_by(year, month) %>% 
  summarize(tmax = mean(tmax)) %>% 
  filter(month == 8) %>% 
  dplyr::select(year, tmax) %>% 
  mutate(tmax = tmax - 273.15) %>% 
  ungroup()

### min temp ###
tmin_8.5_og <- read_csv("../HW_3/data/tasmin_day_CCSM4_rcp85.csv")

# clean and summarize by correct month
tmin_8.5 <- tmin_8.5_og %>% 
  rename(tmin = "tasmin_day_CCSM4_rcp85") %>% 
  mutate(year = year(time)) %>% 
  mutate(month = month(time)) %>% 
  group_by(year, month) %>% 
  summarize(tmin = mean(tmin)) %>% 
  filter(month == 5) %>% 
  dplyr::select(year, tmin) %>% 
  mutate(tmin = tmin - 273.15) %>% 
  ungroup()

### precip ###
precip_8.5_og <- read_csv("../HW_3/data/pr_day_CCSM4_rcp85.csv",
                          col_names=c("time", "precip"), skip = 1)

# clean and summarize by correct month
precip_8.5 <- clean_clim(precip_8.5_og, type = "precip")

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# combine into one dataframe

CCSM4_RCP_8.5 <- left_join(tmax_8.5, tmin_8.5, by = "year") %>% 
  left_join(precip_8.5, by = "year")

# plot temp
ggplot(CCSM4_RCP_8.5, aes(x = year, y = tmax))+
  geom_line() +
  geom_line(aes(y = tmin))
# plot precip
ggplot(CCSM4_RCP_8.5, aes(x = year, y = precip))+
  geom_line() 

# avocado yield anomaly

CCSM4_yield_8.5 <- CCSM4_RCP_8.5 %>% 
  mutate(yield_anomaly = (17.71*lag(tmax,1) - 
                          0.29*lag(tmax,1)^2 + 
                          3.25*tmin - 
                          0.14*tmin^2 + 
                          1.00*lag(precip,1) - 
                          0.31*lag(precip,1)^2 - 
                          288.09))
# plot it
ggplot(CCSM4_yield_8.5, aes(x = year, y = yield_anomaly)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Yield Anomaly (ton/acre)"
  ) +
  theme_classic()
```

CCSM4

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# graph together
CCSM4_yields <- CCSM4_yield_8.5 %>% 
  rename("CCSM4 RCP 8.5" = yield_anomaly) %>% 
  left_join(CCSM4_yield_4.5, by = "year") %>% 
  rename("CCSM4 RCP 4.5" = yield_anomaly) %>% 
  dplyr::select(year, "CCSM4 RCP 4.5", "CCSM4 RCP 8.5") 

CCSM4_gather <- CCSM4_yields %>% 
  gather(key = "projection", value = "anomaly", -year)

# plot
ggplot(CCSM4_gather, aes(x = year, y = anomaly)) +
  geom_line(aes(color = projection)) +
  geom_smooth(aes(color = projection), se = FALSE) +
  labs(
    x = "Year",
    y = "Yield Anomaly (tons/acre)",
    color = "Climate Projections"
  ) + 
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() 

```

All together

```{r}

all_project <- left_join(ACCESS_yields, CCSM4_yields, by = "year") %>%
  gather(key = "projection", value = "anomaly", -year)


# plot
ggplot(all_project, aes(x = year, y = anomaly)) +
  geom_line(aes(color = projection)) +
  geom_smooth(aes(color = projection), 
              size = 1.2, se = FALSE) +
  labs(
    x = "Year",
    y = "Yield Anomaly (tons/acre)",
    color = "Climate Projections"
  ) + 
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(limits = c(-6,2)) +
  scale_color_manual(values = c("lightsalmon4", "lightsalmon3",
                                "darkolivegreen4", "darkolivegreen3")) +
  theme_classic() 

```

```{r}
# plot
ggplot(all_project, aes(x = year, y = anomaly, fill = projection)) +
  geom_smooth(aes(color = projection)) +
  labs(
    x = "Year",
    y = "Yield Anomaly (tons/acre)",
    color = "Climate Projections"
  ) + 
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(limits = c(-6,2)) +
  scale_color_manual(values = c("lightsalmon4", "lightsalmon3",
                                "darkolivegreen4", "darkolivegreen3")) +
  scale_fill_manual(values = c("lightsalmon4", "lightsalmon3",
                                "darkolivegreen4", "darkolivegreen3")) +
  theme_classic() 


```

# moving average
```{r}

# plot
ggplot(all_project, aes(x = year, y = anomaly, fill = projection)) +
  geom_ma(ma_fun = SMA, n = 10, 
          aes(color = projection, linetype = "solid")) + 
  labs(
    x = "Year",
    y = "Yield Anomaly (tons/acre)",
    color = "Climate Projections"
  ) + 
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(limits = c(-6,2)) +
  scale_color_manual(values = c("lightsalmon4", "lightsalmon3",
                                "darkolivegreen4", "darkolivegreen3")) +
  scale_fill_manual(values = c("lightsalmon4", "lightsalmon3",
                                "darkolivegreen4", "darkolivegreen3")) +
  theme_classic() 




```

# averages for 2007-2037 and 2070-2100
```{r}

av_06 <- all_project %>% 
  filter(year %in% c(2007:2036)) %>% 
  group_by(projection) %>% 
  summarize(av_yield = mean(anomaly))

av_70 <- all_project %>% 
  filter(year >= 2070) %>% 
  group_by(projection) %>% 
  summarize(av_yield = mean(anomaly)) 

av_yeilds <- left_join(av_06, av_70, by = "projection") %>% 
  mutate(percent_change = ((av_yield.x - av_yield.y)/av_yield.x)*100)
  

```





